Event-Related Potentials of the Semantically\\ Informed Perception of Unfamiliar Objects

### 2020-11-03 analysis/manuscript.Rmd
rename N1 to N170

  The thesis was handed December 18, 2020 and has been supervised by Prof. Dr. Rasha Abdel Rahman, Humboldt-Universität zu Berlin, Germany, and Prof. Dr. Franziska Süß, Fachhochschule des Mittelstands, Bamberg, Germany.

@@

# Chunk options
knitr::opts_chunk$set(include = FALSE, fig.align = "center", fig.out = "100%", fig.width = 10)

@@

How does our perception of an object change as soon as we discover what it is for? In this study, we presented participants with unfamiliar objects, preceded in half of the cases by valid information about their function (leading to semantically informed perception) and in the other half of the cases by misleading information (leading to naive perception). Capitalizing on the high temporal resolution of event-related potentials (ERPs), we investigated if semantically informed versus naive perception led to differential effects at different stages in the processing hierarchy of visual objects.

Learning semantic knowledge about objects has an influence on how we process them [@abdelrahman2008; @gauthier2003; @gratton2009; @maier2014; @maier2019; @rossion2002; @rossion2004; @tanaka2001; @samaha2018; @weller2019]. Recently, evidence has accumulated that it does so not only at "higher" stages in the visual processing hierarchy, i.e. after "lower" stages of perceptual processing have been completed. It rather seems as though knowledge can also alter early processes, previously thought to be determined solely by the visual input itself, in a top-down manner. Such a top-down influence of semantic knowledge or expertise may manifest itself in behavioral changes such as improved visual matching [e.g.  @gauthier2003] and visual search performance [@maier2014; @holmes2012] when having learned that two objects share the same semantic category or verbal label. Furthermore, acquiring semantically meaningful knowledge about previously unfamiliar objects has been shown to increase their likelihood of gaining access to conuscious visual awareness under circumstances when attentional resources are limited [@weller2019]. Although these behavioral effects hint at an early locus of semantic knowledge effects in object processing, they in and of themselves cannot convincingly demonstrate that not just downstream, postperceptual steps are being affected [@firestone2016].

Stronger evidence comes from some of the above and other studies concurrently measuring event-related brain potentials (ERPs) while participants are performing different tasks on the objects about which they had previously acquired semantic knowledge. In their ERP study, [@abdelrahman2008] had participants first learn in-depth knowledge about 20 unfamiliar objects by listening to a verbal descripition of the function of each object while it was presented visually on the screen, whereas for 20 different objects they did not receive any relevant semantic information, instead listening to unrelated cooking recipes while the objects were presented. As in the present study, the assignment of unfamiliar objects to either of the two conditions was counterbalanced ascross participants as to prevent any subsequent effects of semantic knowledge from being confounded with visual differences between the objects. In a separate test session, event-related potentials were measured while participants performed three different tasks on those objects (intermixed with well known objects that had not been part of the training session). ERP amplitudes in response to objects for which in-depth versus minimal knowledge had been acquired differed in the time window of the N400 component (300--500 ms after stimulus onset), typically associated with semantic processing, but, interestingly, aslo in the time window of the visual P1 component (100--150 ms after stimulus onset), typcially associated with lower-level visual processing [[@abdelrahman2008]; Experiment 1]. This effect ocurred across three different tasks (none of which explicitly required accessing any of the aqcuired information) and was qualitatively similar to the ERPs observed for untrained but well-known objects. A modulation of the P1 component has also been replicated using similar sets of objects in the above mentioned studies showing effects of object knowledge in visual search and the visual awareness tasks [@maier2019; @weller2019]. It thus seems as though learning semantic information can influence steps of the processing hierarchy previously thought to be determined by the low-level visual input.

A different strategy to investaige the influence of semantic knowledge on perception has been taken in a recent ERP study by [@samaha2018]. Instead of using unfamiliar objects as stimuli, participants viewed images of familiar everyday objects that had been converted to two-tone "Mooney" images, making the object itself difficult to recognize. For half of the images, participants performed a semantic training where they were told which object to look for in each image. For the other half of images, they performed a perceptual (1-back) control task to equate the amount familiarity across all images. In a subsequent visual discrimination task, participants tended to respond faster and more accurate to the semantically trained images than to the perceptually trained ones [[@samaha2018]; Experiment 4]. Crucially, the semantic training also manifested itself in increased pre-stimulus alpha-band power and larger P1 amplitudes at left temporo-occipital electrodes, further coroborating an early influence of semantically informed as compared to naive perception of objects.

What all the previous studies have in common is that they investigated the effects of semantic knoweldge on object perception *after* an extensive learning phase [@abdelrahman2008; @gauthier2003; @maier2014; @maier2019; @rossion2002; @rossion2004; @samaha2018; @weller2019]. Participants usually performed a seperate training session where they encountered each object together with its respective label or description multiple times. The EEG was usually not measured (or at least not analyzed and reported) during this training phase, which sometimes took place on a separate day [@abdelrahman2008; @maier2014; @maier2019] or was even spread out across multiple days or weeks [@rossion2002; @rossion2004]. While designs with such substantive learning maximizes the chances of detecting even subtle top-down effects of the acquired knowledge, they necessarily leave at least three important conceptual questions unresolved. The first question is if we can detect any electrophysiological correlates of semantic insight, i.e. the critical presentation during which an understanding of a previously unfamiliar or meaningless objects is happening, instead of asking if this understanding leads to differential effects later in an orthogonal task. The second question is how much learning is actually necessary before reliable top-down effects of knowledge (e.g. on object-related P1 amplitudes) can be obtained: Does it actually take dozens of repitions per object or may a single exposure to the object together with the relevant semantic information be enough? The third and closely related question is related to one of the critiques sometimes raised against the nature of knowledge effects: Are they reflecting genuine top-down effects of semantic processes, altering perception online (while the object is being perceived), or do they merely reflect the (re-)activation of stored representations of the objects which have been created or altered throughout the learning process? By measuring the ERPs before, during, and after participants received semantic hints about the unfamiliar objects, we hope to provide tentative answers to all of these questions.

-   One paragraph explaining that previous research has shown that semantic knowledge modulates early stages of visual perception (e.g. Abdel Rahman & Sommer, 2008; Gauthier et al., 2003; Maier & Abdel Rahman, 2019; Maier et al., 2014; Rossion et al., 2002, 2004; Samaha et al., 2018; Tanaka & Curran, 2001; Weller et al., 2019).

-   Potentially one paragraph describing Abdel Rahman & Sommer (2008) in more detail (participants acquiring knowledge about existing but unfamiliar objects).

-   One paragraph describing Samaha et al. (2018) in more detail (participants being trained to see familiar objects in otherwise meaningless images). Possibly hypothesize a special role of insight (= the moment when understanding is acquired), although this is not directly captured by the study.

-   One paragraph explaining how all of these studies measured ERPs only *after* an extensive learning phase, making it difficult to determine (a) how much learning is necessary for semantic knowledge to influence perception and, related to that, (b) if the modulation of early ERP components reflects genuine online effects of semantic knowledge or rather the activation of altered visual representations (Firestone & Scholl, 2015) (?).

-   One paragraph on theories of top-down/language effects on visual perception (reverse hierarchy theory).

-   Possibly another paragraph on theories of top-down effects on visual perception (label feedback hypothesis) if it fits thematically.

-   Do we need to refer to the role of attention already in the introduction? If so, potentially another paragraph here. I might prefer to leave that for the discussion.

-   Possibly one more paragraph on why ERPs are so suitable to answer this research question, including a brief characterization of the relevant components (P1: lower-level visual perception, N1: higher-level visual perception, N400: semantic processing).

-   The present study: Design of the study with a focus on distinguishing between online effects and altered representations; no specific hypotheses but looking for influences of semantic knowledge on the three ERP components at three different stages (before, during, and after acquisition).

@@

# Check average number of stimuli per condition
(stims_exp1 <- map(data_exp1, function(x) {
  table(x$condition) / 3
}) %>% bind_rows() %>% colMeans())
```

@@

# # Get indices of bad epochs despite ICA
# rej_exp1 = epochs_exp1.drop_log
# rej_exp1[:] = [item for item in rej_exp1 if item != ['IGNORED']]
# rej_exp1 = [i for i in range(len(rej_exp1)) if rej_exp1[i] != []]

# Define function to compute grand averages per condition
def compute_evokeds(epochs):
    evokeds = dict()
    evokeds_dat = dict()
    # Each part becomes a dictionary of conditions
    for pt in ["I", "II", "III"]:
        evokeds[pt] = dict()
        evokeds_dat[pt] = dict()
        # Each condition becomes a list of participants
        for cn in ["Informed", "Naive"]:
            evokeds[pt][cn] = list()
            evokeds_dat[pt][cn] = list()
            # For every participant we average trials separately for all parts and conditions
            for vp in epochs.metadata["participant"].unique():
                query = 'participant == "' + vp + '" and part == "' + pt + '" and condition == "' + cn + '"'
                evokeds[pt][cn].append(epochs[query].average())
            # Compute grand averages across participants for this part and condition
            evokeds[pt][cn] = mne.grand_average(evokeds[pt][cn])
            # Export only the actual data for R
            evokeds_dat[pt][cn] = evokeds[pt][cn].data
    return (evokeds, evokeds_dat)

@@

@@

# Create vectors with sample time points (in ms) and electrodes
tmin <- py$preproc_params$tmin * 1000
tmax <- py$preproc_params$tmax * 1000
sfreq <- py$epochs_exp1$info["sfreq"]
times <- seq(tmin, tmax, 1000 / sfreq)
els <- py$epochs_exp1$ch_names

# Define ERP components of interest
comps <- tibble(
  name = c("P1", "N1", "N400"),
  start = c(100, 150, 400),
  stop = c(150, 200, 700),
  roi = list(
    c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
    c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
    c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
  )
)

@@

(ref:figure-1-caption) Procedure and results of Experiment 1. *(A)* In the pre-insight part (Part I) participants were presented with 120 unfamiliar objects and indicated whether they knew what kind of object they were viewing. In the insight part (Part II), half of these objects were presented with matching keywords (in purple color for illustration), leading to semantically informed perception, and the other half with non-matching keywords (in petrol color for illustration), leading to naive perception. In the post-insight part (Part III), the same objects were presented again without the keywords. *(B)* ERP waveforms and scalp topographies separately for objects with semantically informed versus naive perception in Parts I, II, and III. Semantically informed perception was associated with significantly more negative amplitudes in the N1 component in Part II, significantly less negative amplitudes in the N400 component in Parts II and III, and significantly more positive amplitudes in the P1 component in Part III. Ampl. = amplitude.\newline\**p* \< .05. \*\**p* \< .01. \*\*\**p* \< .001.

```{r exp1-plot, include=TRUE, fig.height=11, fig.cap = "(ref:figure-1-caption)"}
# Manually extract significance levels from contrasts for plotting
stars_exp1 <- list(
  P1 = c(I = NA, II = NA, III = "*"),
  N1 = c(I = NA, II = "*", III = NA),
  N400 = c(I = NA, II = "***", III = "***")
)

# Import grand averaged evoked potentials from Python
evokeds_exp1 <- py$evokeds_dat_exp1

# Retrieve locations for the relevant electrodes from standard montage
fname_montage <- list.files(path = "renv/python", pattern = "easycap-M1.txt", recursive = TRUE, full.names = TRUE)[1]
montage <- py$evokeds_exp1$I$Informed$ch_names %>%
  as_tibble() %>%
  rename(electrode = value) %>%
  left_join(read_tsv(fname_montage) %>% rename(electrode = Site)) %>%
  mutate(
    x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
    y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi))
  )

# Read SVG file with the trial structure (created using MS PowerPoint)
plot_trials <- grImport2::readPicture("analysis/manuscript_files/figure-1a.svg") %>%
  grImport2::pictureGrob(expansion = 0)

# Define function to plot trial structure
plot_trial <- function() {

@@

  colour_informed <- viridisLite::viridis(1, begin = 0.1, end = 0.1)
  colour_naive <- viridisLite::viridis(1, begin = 0.5, end = 0.5)
  trial_I <- ggplot() +
    coord_fixed() +
    theme_void() +
    xlim(c(0, 74)) +
    ylim(c(-5, 68)) +
    geom_text(aes(x = 37, y = 68, label = "Part I"), size = 14 / .pt, family = "Helvetica", fontface = "bold") +
    geom_rect(aes(xmin = 18, xmax = 38, ymin = 13, ymax = 33), color = "black", fill = "white") +
    geom_text(aes(x = 28, y = 23, label = "+"), size = 5, family = "Helvetica", fontface = "bold") +
    geom_text(aes(x = 28, y = 10, label = "0.5 s"), color = "black", size = 10 / .pt, family = "Helvetica") +
    geom_rect(aes(xmin = 36, xmax = 56, ymin = 26, ymax = 46), color = "black", fill = "white") +
    draw_image("analysis/manuscript_files/potato_masher.png", x = 36, y = 26, width = 20, height = 20) +
    geom_text(aes(x = 47, y = 23, label = "max. 3 s"), color = "black", size = 10 / .pt, family = "Helvetica") +
    geom_segment(aes(x = 46, y = 13, xend = 56, yend = 20.222), arrow = arrow(length = unit(0.2, "cm")),
                 color = "black")
  trial_II <- ggplot() +
    coord_fixed() +
    theme_void() +
    xlim(c(0, 74)) +
    ylim(c(-5, 68)) +
    geom_text(aes(x = 37, y = 68, label = "Part II"), size = 14 / .pt, family = "Helvetica", fontface = "bold") +
    geom_rect(aes(xmin = 0, xmax = 20, ymin = 0, ymax = 20), color = "black", fill = "white") +
    geom_text(aes(x = 10, y = 10.5, label = "+"), size = 5, family = "Helvetica", fontface = "bold") +
    geom_text(aes(x = 10, y = -3, label = "0.5 s"), color = "black", size = 10 / .pt, family = "Helvetica") +
    geom_rect(aes(xmin = 18, xmax = 38, ymin = 13, ymax = 33), color = colour_informed, fill = "white") +
    geom_text(aes(x = 28, y = 23, label = "Potato\nmashing"), color = colour_informed, size = 10 / .pt,
              family = "Helvetica") +
    geom_text(aes(x = 28, y = 10, label = "2.5 s"), color = "black", size = 10 / .pt, family = "Helvetica") +
    geom_rect(aes(xmin = 18, xmax = 38, ymin = 39, ymax = 59), color = colour_naive, fill = "white") +
    geom_text(aes(x = 28, y = 49, label = "Message\nmorsing"), color = colour_naive, size = 10 / .pt,
              family = "Helvetica") +
    geom_text(aes(x = 28, y = 36, label = "or"), color = "black", size = 10 / .pt, family = "Helvetica") +
    geom_rect(aes(xmin = 36, xmax = 56, ymin = 26, ymax = 46), color = "black", fill = "white") +
    geom_text(aes(x = 46, y = 34, label = "*"), size = 8, family = "Helvetica") +
    geom_text(aes(x = 46, y = 23, label = "0.5 s"), color = "black", size = 10 / .pt, family = "Helvetica") +
    geom_rect(aes(xmin = 54, xmax = 74, ymin = 39, ymax = 59), color = "black", fill = "white") +
    draw_image("analysis/manuscript_files/potato_masher.png", x = 54, y = 39, width = 20, height = 20) +
    geom_text(aes(x = 65, y = 36, label = "max. 3 s"), color = "black", size = 10 / .pt, family = "Helvetica",
              lineheight = 1) +

>>>>

  The thesis was handed in December 18, 2020 and has been supervised by Prof. Dr. Rasha Abdel Rahman, Humboldt-Universität zu Berlin, Germany, and Prof. Dr. Franziska Süß, Fachhochschule des Mittelstands, Bamberg, Germany.

@@

# Chunk options
knitr::opts_chunk$set(include = FALSE, fig.align = "center", out.width = "100%", fig.width = 10)

@@

How does our perception of an object change as soon as we discover what it is for? In this study, we presented participants with unfamiliar objects, preceded in half of the cases by valid information about their function, leading to semantically informed perception, and in the other half of the cases by misleading information, leading to naive perception. Capitalizing on the high temporal resolution of event-related brain potentials (ERPs), we investigated at which stages in the processing hierarchy semantically informed and naive perception of visual objects differed from one another.

The claim that higher-level cognitive capacities such as knowledge or language can modulate lower-level perceptual processes has been the source of a long-standing debate. This debate revolves around the question whether or not our perception of the (visual) world should be thought of as an encapsulated module being completely determined by the visual input reaching our retina. According to the former view, higher-level mental states and functions are not able to penetrate perception and kick in only after lower-level perceptual analysis of the visual input has taken pace [@firestone2016; @fodor1983; @fodor1984; @machery2015; @pylyshyn1999]. Conversely, this view is challenged by theoretical and empirical work suggesting that perceptual processing dynamically interacts with other aspects of cognition from early on [@churchland1988; @churchland1994; @collins2014; @lupyan2012; @lupyan2015; @lupyan2020; @thierry2016; @vetter2014]. These aspects may include transient states such as emotions [e.g.  @phelps2016; @bocanegra2009] or intentions [@balcetis2010; @cole2012] as well as learned capacities such as language [e.g.  @boutonnet2015; @maier2018; @mo2011] or declarative knowledge about faces [e.g.  @abdelrahman2011; @suess2015; @eiserbeck2020] and objects [e.g.  @abdelrahman2008; @lin1997; @weller2019]. Despite this wealth of empirical evidence, critics of the idea that cognitive functions can penetrate perception have remained sceptical, pointing out important methodological shortcomings in large parts of this literature [@firestone2016].

One productive way of investigating top-down effects of cognition on perception while addressing some of these concerns is through neurophysiological recording methods such as ERPs: Their high temporal resolution allows for a direct test of which processing stages---perceptual and/or postperceptual---are influenced by higher-level capacities such as semantic knowledge. This is because when participants are presented with visual stimuli, their averaged brain responses show characteristic deflections (ERP components), each of them with its typical polarity (positive or negative), peak latency, spatial scalp distribution, and functional role(s).

The visual P1 component refers to a positive deflection peaking approximately 100--150 ms after stimulus onset at occipital channels. It is generated by multiple sources in the extrastriatal cortices of the middle occipital and fusiform gyri [@dirusso2001; @foxe2002; @mangun1995]. The P1 is mostly thought to reflect the processing of low-level visual characteristics of the stimulus [such as size, luminance, and contrast; @dirusso2001; @luck2014; @johannes1995], which is why it being modulated by higher-level cognitive capacities would challenge a purely bottom-up, encapsulated conception of visual perception. It should be noted, however, that the P1 is reliably modulated by spatial [@luck2000; @mangun1995; @mangun1991] and non-spatial [@taylor2002] attention.

The visual N170 (or N1) component refers to the negative deflection following the P1 and is maximal around 150--200 ms after stimulus onset at occipito-temporal channels. While still influenced by visual parameters of the stimulus [@johannes1995], it is also sensitive to selective attention, at least when participants are required to discriminate between different visual stimuli [@vogel2000; @mangun1991]. Importantly, the N170 is enlarged (i.e. more negative) in response to faces [@bentin1996; @rossion2011] and other visual stimuli for which one happens to be an expert [e.g. birds, dogs, or cars; @gauthier2003a; @tanaka2001]. This category selectivity seems to reflect both the processing of individual visual features of the respective objects (including faces) as well as their configuration [@eimer2011; @jacques2010; @rossion1999; @sagiv2001]. Thus, a modulation of the N170 by cognitive capacities such as knowledge would suggest these factors having an influence on higher-level, holistic perception of the visual stimuli.

Finally, the N400 component refers to more negative ERP amplitudes when stimuli require more as compared to less resources for semantic processing or integration [@kutas2011; @lau2008; @rabovsky2018], starting approximately 300 ms after stimulus onset and being most pronounced at centro-parietal channels. A modulation of this component by cognitive capapacities such as knowedge can be taken as evidence for a post-perceptual influence on stimulus processing.

Previous studies have focused on these and other ERP components to investiage which processing stages are indeed influenced by knowledge or expertise in the domain of visual objects [@abdelrahman2008; @gratton2009; @maier2014; @maier2019; @rossion2002; @rossion2004; @tanaka2001; @samaha2018; @weller2019]. Together, they suggest that not only higher-level semantic ERPs like the N400 component are influenced by acquiring semantic information about objects, but also early components that typcially reflect bottom-up visual processing. This was the case, for instance, when participants were presented with a range of unfamiliar objects, receiving in-depth verbal descriptions about their function for half of the objects and irrelevant verbal information (cooking recipes) for the other half of the objects [@abdelrahman2008, Experiment 1]. After this learning phase, the same objects were presented in three different ERP tasks which did not require explicit access to any of the learned semantic information. In all three tasks, ERP amplitudes in response to objects for which in-depth knowledge had been acquired differed from those in response to objects for which this has not been the case. Crucially, these differences occurred not just in the N400 component, indexing a modulation of post-perceptual semantic processing, but also in the P1 component, suggesting a top-down modulation of lower-level perceptual processing. Of note, the ERPs in the in-depth knowledge condition were qualitatively similar to those for untrained but well-known objects. The modulation of the P1 component was also replicated when the same objects were presented under circumstances of limited attentional resources in an attentional blink paradigm [@weller2019]. Of note, the (neurophysiological) differences in P1 amplitudes between objects with in-depth versus minimal knowledge were correlated with (behavioral) differences in their detection rate during the attentional blink, which can be taken as further evidence of an influence of semantic knowledge on the early visual perception of unfamiliar objects.

A complementary approach has been taken in a recent study investigating the ERPs in response to familiar (rather than unfamiliar) objects, which were rendered difficult to recognize by converting them to two-tone, "Mooney" images [@samaha2018, Experiment 4]. Here, participants were trained with meaningful verbal cues, telling them what kind of objects they should look for in the image beforehand, or with a non-meaningful perceptual task, familiarizing them with the images but not with their semantic content. When the EEG was measured in a separate delayed matching task later on, the two types of training led to differential effects, again suggesting a modulation of lower-level visual perception: Over left posterior electrodes, P1 amplitudes were larger and alpha power was higher for meaning-trained images of objects than for perceptually-trained images of objects. This alternative approach further coroborates an early influence of semantic information on visual object perception.

All the previously mentioned studies have in common that they investigated the effects of semantic knoweldge on object perception only *after* an extensive learning phase [@abdelrahman2008; @gauthier2003; @maier2014; @maier2019; @rossion2002; @rossion2004; @samaha2018; @weller2019]. Participants performed one or multiple training session where they encountered each object together with its respective label or description multiple times. The EEG was usually not measured---or at least not analyzed and reported---during these training sessions, which sometimes took place on a separate day [@abdelrahman2008; @maier2014; @maier2019] or were spread out across multiple days or weeks [@rossion2002; @rossion2004]. While designs with such substantive training maximize the chances of detecting even subtle top-down effects of the knowledge acquired, they necessarily leave at least three important conceptual questions unresolved.

The first question is if we can detect any electrophysiological correlates of semantic insight, i.e. the critical presentation during which an understanding of the previously unfamiliar object is happening, instead of asking if this understanding leads to differential effects in an orthogonal task later on. The second question is how much learning is actually necessary before reliable top-down effects of knowledge (such as on object-related P1 amplitudes) can be obtained: Does it actually take dozens of repitions per object or may a single exposure to the object together with the relevant semantic information be enough? The third and closely related question is about the nature of knowledge effects on perception: Are they reflecting genuine top-down effects of semantics, altering perception online (while the object is being perceived), or do they merely reflect the (re-)activation of stored representations of the objects, having been altered throughout the learning process [@haber1966]? By measuring the ERPs before, during, and after participants received semantic hints about unfamiliar objects, we hope to provide tentative answers to all three of these questions.

In the two experiments subsequently reported, participants were presented with real-word objects that were presumed to be unfamiliar to most of them. They first viewed each of these objects once without any semantic information. This part served as a naive baseline to outrule that ERPs would differ in response to the objects based solely on their low-level visual differences. Next, participants viewed each unfamiliar object for a second time, now preceded by verbal keywords. These keywords could either correctly describe the object's typical function, thus making it possible for participants to understand what kind of object they are viewing, or mislead them by describing the function of a different object, thus keeping the perception of the object semantically naive. During this "insight" part, we were able measure the online influence of semantic knowledge on object-evoked ERPs as it happened. Finally, the objects were presented for a third time (without keywords, as in the first part) to investigate downstream effects of having acquired semantic knowledge about them, thus mimicking previous studies [e.g. @abdelrahman2008, @samaha2018; @weller2019]. In all three parts, we compared the influences of semantic information on ERPs associated with lower-level visual perception (P1 component, 100--150 ms), higher-level visual perception (N170 component, 150--200 ms), and semantic processing (400--700 ms).

@@

# Check average number of stimuli per condition
(stims_exp1 <- map(data_exp1, function(x) {table(x$condition) / 3}) %>% bind_rows() %>% colMeans())
```

```{r}
# vec_participant <- data_exp1 %>% bind_rows() %>% pull(participant) %>% unique()
# vec_participant <- c(vec_participant[seq(1, 23, 2)], vec_participant[seq(2, 24, 2)])
# 
# data_exp1 %>%
#   bind_rows() %>%
#   mutate(participant = factor(participant, levels = vec_participant),
#          item_new = ifelse(participant %in% vec_participant[1:12], item - 15, item)) %>%
#   filter(part == "II" & Tastencode != 0) %>%
#   ggplot(aes(x = participant, y = item, fill = Tastencode)) +
#   geom_raster() +
#   scale_fill_viridis_c(begin = 0.1, end = 0.5)
#   #scale_fill_manual(values = c("Informed" = "purple", "Naive" = "green", "Excl_informed" = "gray70", "Excl_naive" = "gray90", "Excl_known" = "white"))
```

@@

# Define function to compute grand averages per condition
def compute_evokeds(epochs):
    evokeds = dict(); evokeds_dat = dict()
    # Each part becomes a dictionary of conditions
    for pt in ["I", "II", "III"]:
        evokeds[pt] = dict(); evokeds_dat[pt] = dict()
        # Each condition becomes a list of participants
        for cn in ["Informed", "Naive"]:
            evokeds[pt][cn] = list(); evokeds_dat[pt][cn] = list()
            # For every participant we average trials separately for all parts and conditions
            for vp in epochs.metadata["participant"].unique():
                query = 'participant == "' + vp + '" and part == "' + pt + '" and condition == "' + cn + '"'
                evokeds[pt][cn].append(epochs[query].average())
            # Compute grand averages across participants for this part and condition
            evokeds[pt][cn] = mne.grand_average(evokeds[pt][cn])
            # Export only the actual data for R
            evokeds_dat[pt][cn] = evokeds[pt][cn].data
    return (evokeds, evokeds_dat)

@@

```{r exp1-tryout, eval=FALSE}
# # Create vector of sample time points
# times <- seq(py$preproc_params$tmin, py$preproc_params$tmax, 1 / py$epochs_exp1$info["sfreq"]) * 1000
# 
# # Average channels in the N170 ROI
# py_run_string('epochs_roi = mne.channels.combine_channels(epochs_exp1, groups=dict(roi=mne.pick_channels(epochs_exp1.ch_names, ["P7", "P8", "PO7", "PO8", "PO9", "PO10"])))')
# 
# # Extract data for Part II from Python
# dat <- py$epochs_exp1$metadata %>%
#   as_tibble() %>%
#   bind_cols(py$epochs_roi$get_data(picks = "roi") %>% drop() %>% as_tibble()) %>%
#   pivot_longer(-c(part, condition, participant, item, RT), values_to = "amplitude") %>%
#   filter(condition %in% c("Informed", "Naive")) %>%
#   mutate(
#     part(factor(part)),
#     condition = factor(condition),
#     participant = factor(participant),
#     item = factor(item),
#     time = rep(times, nrow(.) / length(times))
#   )
# 
# # Compute LMM
# library(lme4)
# contrasts(dat$condition) <- contr.sum(2)
# contrasts(dat$part) <- MASS::contr.sdif(3)
# mod <- lmer(
#   amplitude ~ part/condition + (part/condition|time:participant) + (part/condition|time:item),
#   data = dat, control = lmerControl(calc.derivs = FALSE)
# )
# summary(mod)
# 
# # Predict values from LMM
# probe <- expand.grid(
#   part = unique(dat$part),
#   condition = unique(dat$condition),
#   participant = unique(dat$participant),
#   item = unique(dat$item),
#   time = dat$time %>% unique()
# ) %>% mutate(amplitude = predict(mod, .))
# 
# # Plot real and modelled data
# ggplot(mapping = aes(x = time, y = amplitude, color = condition)) +
#   geom_line(data = dat, aes(group = interaction(condition, participant, item)), size = 0.1, alpha = 0.1) +
#   stat_summary(data = probe, geom = "line", fun = mean, na.rm = TRUE, size = 1) +
#   theme_minimal() +
#   labs(x = "Time (ms)", y = "Predicted amplitude (µV)") +
#   coord_cartesian(xlim = c(-200, 800), ylim = c(-15, 15), expand = FALSE)
# 
# # Old models
# mod1_P1 <- lmerTest::lmer(P1 ~ part/condition + (part/condition|participant) + (part/condition|item),
#                           data = data_exp1, control = ctrl_params)
# mod1_N170 <- lmerTest::lmer(N170 ~ part/condition + (part/condition|participant) + (part/condition|item),
#                           data = data_exp1, control = ctrl_params)
# mod1_N400 <- lmerTest::lmer(N400 ~ part/condition + (part/condition|participant) + (part/condition|item),
#                             data = data_exp1, control = ctrl_params)
# mod2_P1 <- lmerTest::lmer(P1 ~ part/condition + (part/condition|participant) + (part/condition|item),
#                           data = data_exp2, control = ctrl_params)
# mod2_N170 <- lmerTest::lmer(N170 ~ part/condition + (part/condition|participant) + (part/condition|item),
#                           data = data_exp2, control = ctrl_params)
# mod2_N400 <- lmerTest::lmer(N400 ~ part/condition + (part/condition|participant) + (part/condition|item),
#                             data = data_exp2, control = ctrl_params)
# summary(mod1_P1)
# summary(mod1_N170)
# summary(mod1_N400)
# summary(mod2_P1)
# summary(mod2_N170)
# summary(mod2_N400)
```

@@

# Create vectors with sample time points (in ms) and electrodes
times <- seq(py$preproc_params$tmin * 1000, py$preproc_params$tmax * 1000, 1 / py$epochs_exp1$info["sfreq"] * 1000)
els <- py$epochs_exp1$ch_names

# Define ERP components of interest
comps <- tibble(
  name = c("P1", "N170", "N400"),
  start = c(100, 150, 400),
  stop = c(150, 200, 700),
  roi = list(
    c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
    c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
    c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
  )
)

@@

(ref:figure-1-caption) Procedure and results of Experiment 1. *(A)* In the pre-insight part (Part I), participants were presented with 120 unfamiliar objects and indicated whether they knew what kind of object they were viewing. In the insight part (Part II), half of these objects were presented with matching keywords (in purple color for illustration), leading to semantically informed perception, and the other half with non-matching keywords (in petrol color for illustration), leading to naive perception. In the post-insight part (Part III), the same objects were presented again without the keywords. *(B)* ERP waveforms and scalp topographies are shown for objects with semantically informed versus naive perception within Parts I, II, and III. Semantically informed perception was associated with significantly more negative amplitudes in the N170 component in Part II, significantly less negative amplitudes in the N400 component in Parts II and III, and significantly more positive amplitudes in the P1 component in Part III. Ampl. = amplitude.\newline\**p* \< .05. \*\**p* \< .01. \*\*\**p* \< .001.

```{r exp1-plot, include=TRUE, fig.height=11, fig.cap = "(ref:figure-1-caption)"}
# Retrieve locations for the relevant electrodes from standard montage
fname_montage <- list.files(path = "renv/python", pattern = "easycap-M1.txt", recursive = TRUE, full.names = TRUE)[1]
montage <- py$evokeds_exp1$I$Informed$ch_names %>%
  as_tibble() %>%
  rename(electrode = value) %>%
  left_join(read_tsv(fname_montage) %>% rename(electrode = Site), by = "electrode") %>%
  mutate(
    x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
    y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi))
  )

# Trial structure
plot_trials <- function(){

@@

  color_informed <- viridisLite::viridis(1, begin = 0.1)
  color_naive <- viridisLite::viridis(1, begin = 0.5)
  ggplot() +
    coord_cartesian(xlim = c(-100, 100), ylim = c(0, 100), expand = FALSE) +
    ## PART II ##
    # Title
    annotate("text", x = 0, y = 97.2, label = "Part II (insight)", size = 14 / .pt, family = "Helvetica", fontface = "bold") +
    # Arrow
    annotate("segment", x = -30, xend = 24, y = 79, yend = 85, arrow = arrow(length = unit(0.1, "inches"))) +
    # Squares
    annotate("rect", xmin = -37, xmax = -17, ymin = 55, ymax = 75, color = "black", fill = "white") +
    annotate("rect", xmin = -19, xmax = 1, ymin = 45, ymax = 65, color = color_informed, fill = "white") +
    annotate("rect", xmin = -19, xmax = 1, ymin = 69, ymax = 89, color = color_naive, fill = "white") +
    annotate("rect", xmin = -1, xmax = 19, ymin = 59, ymax = 79, color = "black", fill = "white") +
    annotate("rect", xmin = 17, xmax = 37, ymin = 61, ymax = 81, color = "black", fill = "white") +
    annotate("text", x = -27, y = 65.3, label = "+", size = 20 / .pt, family = "Helvetica") +
    annotate("text", x = -9, y = 55, label = "Potato\nmashing", size = 10 / .pt, family = "Helvetica", color = color_informed) +
    annotate("text", x = -9, y = 79, label = "Message\nmorsing", size = 10 / .pt, family = "Helvetica", color = color_naive) +
    annotate("text", x = 9, y = 67, label = "*", size = 30 / .pt, family = "Helvetica") +
    draw_image("analysis/manuscript_files/potato_masher.png", x = 18, y = 62, width = 18, height = 18) +
    # Timings
    annotate("text", x = -9, y = 67.3, label = "or", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = -27, y = 52, label = "0.5 s", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = -9, y = 42, label = "2.5 s", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = 9, y = 56, label = "0.5 s", size = 10 / .pt, family = "Helvetica") +
    annotate("text", x = 27, y = 56.3, label = "3 s or\nresponse", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    # Response options
    annotate("segment", x = 27, y = 39, xend = 27, yend = 51.5) +
    annotate("segment", x = 0, y = 39, xend = 27, yend = 39) +
    annotate("segment", x = 0, y = 35, xend = 0, yend = 39) +
    annotate("segment", x = -32, y = 35, xend = 32, yend = 35) +
    annotate("segment", x = -32, y = 30.5, xend = -32, yend = 35) +
    annotate("segment", x = -12, y = 32, xend = -12, yend = 35) +
    annotate("segment", x = 12, y = 32, xend = 12, yend = 35) +
    annotate("segment", x = 32, y = 32, xend = 32, yend = 35) +
    annotate("text", x = -32, y = 26, label = "A. \"I know\nwhat this is\"", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    annotate("text", x = -12, y = 26, label = "B. \"I have an\nassumption\nwhat this is\"", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
    annotate("text", x = 12, y = 26, label = "C. \"I have rather\nno assumption\nwhat this is\"", size = 10 / .pt, family = "Helvetica", lineheight = 1) +
